---
title: "Hw6"
author: "Getong Zhong"
date: "2022-10-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggExtra)
```

## Part I
### Problem 1

#### (a)
```{r}

dec <- read.csv("C:/Users/tonyg/Desktop/Academic/Grad/HUDM 5026/decath.csv")
is.tibble(dec)
```
It is not tibble since it print out more than 10 rows 

#### (b)
```{r}
print(dec, max.levels = NULL, width = 300)
```


### Problem 2
```{r}
apply(X = dec, MARGIN = 2, FUN = mean)
apply(X = dec, MARGIN = 2, FUN = sd)
```

### Problem 3
```{r}
cov(dec)
```

```{r}
round(cor(dec),2)
```
From the correlation matrice above we can see for X100 it has a pairwise correlation with long, X400, and X110 for a absolute value over 0.5, for long it has a high correlation when pair with long X100 and X400 as well. I think these pairwise correlation kind of make sense to me. Because X100, X400 amd X110 are all mid and short distance so it is not suprise that they have a strong positive corelation with each other, and for long jump it is a totally different sport so it is not suprise that it has a strong negative correlation with the speeding games. 
### Problem 4
```{r}
library(corrplot)
corrplot(cor(dec), order = "hclust") 
```
Disq (discus throw) and poid (shotput) are pretty similar games which require athelet to have a great strength, and even this the similar rules that the are competing how far you can throw/push the heavy ball/discus , so it is make sense that they are inter-correlated .

### Problem 5
#### (a)
```{r}
plot(dec$disq, dec$poid, main = "discus vs shot put (R base)",
     xlab = "Scores for discus", 
     ylab = "Scores for shot put")

```

#### (b)
```{r}
plot(dec$poid, dec$jave, main = "shot put vs javelin (R base)",
     xlab = "Scores for shot put", 
     ylab = "Scores for javelin")
```

#### (c)
```{r}
plot(dec$disq, dec$jave, main = "discus vs javelin (R base)",
     xlab = "Scores for discus", 
     ylab = "Scores for javelin")
```

#### (d)
```{r}
hist(dec$poid, main = "shot put Histogram (R base)",
     xlab = "Scores for shot put", 
     ylab = "Count")
```
### Problem 6
```{r}
par(mfrow = c(2,2))
plot(dec$disq, dec$poid, main = "discus vs shot put (R base)",
     xlab = "Scores for discus", 
     ylab = "Scores for shot put")
plot(dec$poid, dec$jave, main = "shot put vs javelin (R base)",
     xlab = "Scores for shot put", 
     ylab = "Scores for javelin")
plot(dec$disq, dec$jave, main = "discus vs javelin (R base)",
     xlab = "Scores for discus", 
     ylab = "Scores for javelin")
hist(dec$poid, main = "shot put Histogram (R base)",
     xlab = "Scores for shot put", 
     ylab = "Count")
```

### Problem 7
```{r}
p1 <- ggplot(data = dec, mapping = aes(x = disq, y = poid)) +
  geom_point() +
  xlab("Scores for discus") + 
  ylab("Scores for shot put") +
  theme(legend.position = "none")
  

ggMarginal(p1, type = "density")
```

## Part II
### Problem 1
```{r}
acu <- read.csv("C:/Users/tonyg/Desktop/Academic/Grad/HUDM 5026/acupuncture.csv", header = TRUE)

```

### Problem 2
```{r}
acu2 <- acu[
  order(acu$group, acu$sex, acu$age),
]

```

### Problem 3
```{r}
acu3 <- subset(acu, group == 1 & age > 30)

```

### Problem 4
```{r}
fourty_percentile <- quantile(acu$pk1, c(0.4))

acu4 <- subset(acu, pk1 >= fourty_percentile)


```

### Problem 5
```{r}
acu[which(acu$sex == 1 & acu$age >= 40), ]
```